---
title: "class11-"
format: pdf 
Author: 
--- 

**Background**
We saw last day that the main repository for biomolecular structure (the PDB database) only has ~250,000 entries. UniProtKB (the main protein sequence database) has over 200 million entries!


**AlphaFold** 

In this hands-on session we will utilize AlphaFold to predict protein structure from sequence (Jumper et al. 2021). 

Without the aid of such approaches, it can take years of expensive laboratory work to determine the structure of just one protein. With AlphaFold we can now accurately compute a typical protein structure in as little as ten minutes. 

## The EBI AlphaFold Database

The EBI AlphaFold database contains lots of computed structure models. It is increasingly likely that the structure you are interested in is already in this database  
(<https://alphafold.ebi.ac.uk/>).

There are three major outputs from AlphaFold:
1. A predicted protein structure in PDB format  
2. A pLDDT score, which indicates how confident the model is for each residue (values above 70 indicate high confidence)  
3. A PAE score, which provides information about protein packing quality  

If you cannot find a matching entry for your sequence in the AlphaFold Database, you can run AlphaFold yourself.

## Running AlphaFold

In this lab, we use ColabFold to run AlphaFold on our sequence.

- https://github.com/sokrypton/ColabFold  
- https://colab.research.google.com/github/sokrypton/ColabFold/blob/main/AlphaFold2.ipynb  

Figure from AlphaFold shown below.
 
 
```{r}
library(png)
library(grid)

img <- readPNG("class11p.png")
grid.newpage()
grid.raster(img)
```
```{r}
img2 <- readPNG("classp2.png")
grid.newpage()
grid.raster(img2)
```
**Interpreting results** 
Custom analysis of resulting models We can read all the AlphaFold results into R and do more quantitative analysis than just viewing the structures in Mol-star:

Read all PDB models:
```{r}
library(bio3d)

results_dir <- "hivprdimer_23119"
Pdb_file <-  list.files(path = results_dir,pattern = "*.pdb", full.names = TRUE)
print(Pdb_file)
```
```{r}
Pdb_file <- list.files(results_dir, pattern="\\.pdb$", full.names=TRUE)
print(Pdb_file)

```
```{r}
results_dir <- "hivpr_23119"

```

```{r}
list.files(results_dir)

```

```{r}
Pdb_file <- list.files(results_dir, pattern="\\.pdb$", full.names=TRUE)
print(Pdb_file)
```
```{r}

library(bio3d)
library(msa)
library(Biostrings)
pdbs <- pdbaln(Pdb_file, fit = TRUE, exefile = "msa")
```


```{r}
rd <- rmsd(pdbs, fit = TRUE)
rd
```

```



