---
title: "Class 5: Data Viz with ggplot"
author: "Zahra Rashidi (PID A18561538)"
format: pdf
---

## Background

there are lots of ways to make figures in R. These include so-called "base R" graphics (e.g. 'plot()') and tones off add-on packages like **ggplot2**.

For example here we make the same plot with both:
```{r}
head(cars)
```
First, I need to install the package using `install.packages()`.
install.packages("ggplot2") 
N.B. We never run on install cmd in a quarto code chunk or we will end up re-installing packages many many times-which is not what we want! 
  
Everytime we want to use of these “add-on” packages we need to load it up in R with the library() function: 
```{r} 

library(ggplot2)

head(cars)
```

```{r}
plot(cars)
```

#ggplot(cars) 

every ggplot needs at least 3 things:
- The **data**, the stuff you want plotted 
- The **aes** thitics, how the data map to the plot 
-The **geom**etry, the type of plot
```{r}
p <- ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point() + geom_smooth(method="lm",se=FALSE) + labs(title="Stopping distance of old cars", subtitle = "Data from the 'cars' object", x="Speed (mph)", y="Distamce (ft)")
p
```

##Gene expression plot 
 we can read the input data from the class website
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
```{r}
ggplot(genes)+ aes(Condition1, Condition2)+ geom_point()
```
```{r}
table(genes$State)
```
 Version 2 let’s color by State so we can see the “up” and “down” significant genes compared to all the unchanging genes 
```{r}
ggplot(genes)+ aes(Condition1, Condition2, col=State)+ geom_point()
```
 Version 3 plots, let’s modify the default colors to something we like 
```{r}
ggplot(genes)+ aes(Condition1, Condition2, col=State)+ geom_point() + scale_colour_manual( values=c("red", "purple", "darkgreen"))
```
```{r}
ggplot(genes)+
  aes(Condition1, Condition2, col=State)+
  geom_point() +
  scale_colour_manual(values=c("red", "blue", "magenta")) +
  labs(
    x="Control (no drugs)",
    y="Drug",
    title="Gene Expression Changes upon GLP-1 drug"
  ) +
  theme_bw()
```
 
```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"
gapminder <- read.delim(url)
```

```{r}
head(gapminder, 3)
```
```{r}
ggplot(gapminder)+
  aes(gdpPercap, lifeExp)+
  geom_point(alpha=0.3, col="blue")
```
```{r}
ggplot(gapminder)+
  aes(gdpPercap, lifeExp,col=continent)+
  geom_point(alpha=0.3)

```
```{r}
ggplot(gapminder)+
  aes(gdpPercap, lifeExp,col=continent)+
  geom_point(alpha=0.3)+ 
  facet_wrap(~continent)
```


##Custon plots 
how big is this gapmider dataset? 

```{r}
nrow(gapminder)
```
I want to "filter" down to a subset of this data. I will use the **dplyr** package to help me.
First I need to install it and then load it up.....  
install. packages("₫plyr") and then 'library(dyplr)' 
```{r}
library(dplyr)
``` 
trying URL
"https://cran.rstudio.com/bin/macosx/big-sur-arm64/contrib/4.5/dplyr_1.1.4.tgz'
Content type 'application/x-gzip' length 1614595 bytes (1.5 MB)
```{r}
library(dplyr) 
gapminder_2007 <- filter(gapminder, year==2007)
head(gapminder_2007)
```

```{r}
filter(gapminder, year==2007, country=="United States")
```
>Q Make a plot comparing 1977 and 2007 for all countries

```{r}
input <- filter(gapminder, year %in% c(1977, 2007))
head(input)
```
```{r}
input <- filter(gapminder, year %in% c(1977,2007))

ggplot(data=input, aes(gdpPercap, lifeExp, col=continent)) +
geom_point() + facet_wrap(~year)
```



