---
title: "Class 7: Machine Learning"
author: "Zahra Rashidi (PID:"
format: "pdf"
---

## Background

Today we will begin our exloration of important machine learning methods with a focus on **clustering** and **dimensionallity reduction**.

To start testing these methods let's make up some sample data to cluster where we know what the answer should be.  

```{r}
hi
```
```{r}
hist(rnorm(3000, mean=10))
```


>Q. Can you generate 30 numbers centered at +3 and 30 numbers at -3 
taken random from a normal distribution?  


```{r}
 tmp<- c(rnorm(30, mean=3),
      rnorm(30, mean=-3))  

x<- cbind(x=tmp,y=rev(tmp)) 
plot(x)
```
## K-means clustering 

The main function in "base R"for K-means clustering is called 'kmeans()', let's try it out: 
```{r}
k<-kmeans(x, centers =2)

```
>Q. What component of your kmeans results object has the cluster centers?  

```{r}
k$centers
```
>Q. what components of your kmeans result object has the cluster membership vector (i.e the main clustering results: which points are in which cluster)?  

```{r}
k$size 
```

>Q.Plot the results of clustering (i.e our data colored by the clustering results )along with the clustering centers.


```{r}
plot (x, col=k$cluster) 
points(k$centers, col="blue", pch=15, cex=2)
```
>Q. Can you run 'kmeans()' again and cluster 'x' into 4 clusters and plot the results just like did above with coloring by cluster and the cluster centers shown un blue? 

```{r}
k4 <- kmeans(x, centers = 4)

plot(x, col = k4$cluster)
points(k4$centers, col = "blue", pch = 19, cex = 2)
```

>** Key-point:** Kmeans will always return the clustering that we ask for (this is the "k" or "centers" in K-means)! 

```{r}
k$tot.withinss
```

## Hierechical clustering  

the main function for hierarchical clustering in base R is called 'hclust()'. one of the mian differnces with respect to the base 'kmeans()' function is that you can not just pass your input data directly to 'hclust()' - it needs a "distnace matrix" as input. We can get this from lot's of places including the 'dist()' function. 

```{r}
d <-dist (x) 
hc<- hclust(d)
plot (hc)
```
We can cut dendrogram or "tree" given height to yield our "clusters". For this we use the funtion 'cutree'

```{r}
plot(hc)
     abline(h=10, col="red") 
     cutree(hc, h=10) 
grps <- cutree(hc, h = 10)
```
```{r}
grps
```
>Q. Plot our data 'x' colored by the clustering resul from 'hclust()' and 'cutree'? 


```{r}
grps <- cutree(hc, h = 10) 
plot(x, col=grps)
```


```{r}
plot(hc)
     abline(h=4.2, col="red") 
     cutree(hc, h=4.2) 
```
## Principal Component Analysis (PCA)

PCA is a popular dimensionality reduction technique that is widely used in bioinformatics. 


### PCA of UK food data 

start the lab sheet 

Read data on food consumption in the UK 


```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
```
it looks like the row names are not set properly. We can fix this

```{r}
rownames(x) <- x[,1]
x <- x[, -1]
x

```

A better way to do this is fix the raw names assignment at import time: 
```{r}
x <- read.csv(url, row.names = 1)

```

> Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions? 

```{r}

```


> Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?   

```{r}
# Using base R
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))

```


> Q3: Changing what optional argument in the above barplot() function results in the following plot? 


```{r}
# Currently we have wide format
dim(x)

```



> Q4: missing

> Q5: We can use the pairs() function to generate all pairwise plots for our countries. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?


```{r}
pairs(x, col=rainbow(nrow(x)), pch=16)
```

## Heatmap 

We can install the **pheatmap** package with the ' install.packages()' command that we used previously. Remember that we always run thid in the cansole 

Of all of these plot really only the 'pairs()' plot was useful. This however took a bit of work to interpret and will at scale when I am looking at much bigger datasets.  


## PCA the rescue 
The main function in "base R" for PCA is called 'prcamp()'. 

```{r}
pca <- prcamp( t(x) ) 
summary(pca)
```

>Q. how much variance is captured in the first PC? 
67.4% 
>Q. how many PC's do i need to capture at leadt 90% of the total varance in the dataset? 

Two Pc's captutr 96.5% of the total variance. 

>Q plot our main PCA results. Folks can call this different thing depending on their filed of study e.g "PC plot", "Ordienation plot" "Score Plot", 

```{r}
pca$x
```
 
 
 ## Digging deeper (variable loadings)
 
 How do the orginional variables (i.e the 17 differen food) 
 contribute to our new PCs
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
